---
title: "2 - Citation of Research Outputs"
date: "`r Sys.Date()`"
always_allow_html: yes
output:
  md_document:
    variant: markdown_github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = FALSE)
library(dplyr);library(impactr)
data_presentations <- readRDS("~/starsurg_impact/data/data_pre.rds") %>%
  dplyr::filter(level == "International") %>%
  dplyr::distinct(title, .keep_all = TRUE) %>%
  dplyr::mutate(title = ifelse(title=="Collaborative Research: How to Get Involved", NA, title)) %>%
  dplyr::mutate(author = ifelse(con_city=="Belfast", NA, author)) %>%
  dplyr::select(type:author) %>% dplyr::arrange(con_date_start) %>% tail(5)

out_doi <- impactr::extract_doi(doi = c("10.1136/bmjopen-2015-009812", "10.1136/bmjopen-2015-008811", "10.1002/bjs5.86"),
                       get_auth = TRUE, get_almetric = FALSE, get_impact = FALSE)

```

# **Citation of Research Outputs**
The most common vehicles for research impact tend to be for presentations and publications of the work conducted. These often are often listed at both an individual- (CV) and institutional-level to demonstrate scholary activity.

## **1. `cite_publication()`**
The `cite_publication()` function will accept direct input from both `extract_pmid()` and `extract_doi()`, however will also accept other dataframes with the prerequiste columns. These columns can be specified within the function. Other features include:

- Highlighting any missing data essential for citation (e.g. author, jounral, etc). The issue, pmid, and doi are considered optional (depending on whether the dataset was generated from `extract_pmid()` or `extract_doi()`).

- Automatically displaying all authors, however the auth_max can be set from 1 to n to add "et al." for any authors beyond that number.

- The format can be customised via `cite_format` to match preferred referencing style (default is as shown below - Vancouver). Note the variable names in the string must exactly match an essential column (e.g. "author" not "Author" or "AUTHOR") 

```{r edit, warning=FALSE, message=FALSE}
 out_doi %>%
  # If a single-authorship collaborative publication (e.g. "STARSurg Collaborative") then display that
  dplyr::mutate(author = ifelse(is.na(author_group)==TRUE, author, author_group)) %>%
  
  impactr::cite_publication(journal = "journal_full", max_auth = 10,
                   cite_format = "author. title. journal. year; volume (issue): pages. PMID: pmid. DOI: doi.") %>%
  
  dplyr::mutate(citation = gsub("\\]", "", gsub("\\[", "", as.character(citation)))) %>%
  dplyr::select(citation) %>%
  knitr::kable(format="html") %>% kableExtra::kable_styling(bootstrap_options = "striped", full_width = F)
  
```

&nbsp;

## **2. `cite_presentation()`**
Presentations of academic work are important additional research outputs, yet are often not recorded online and so cannot be extracted. Therefore, the `cite_presentation()` function will accept any dataframe with the prerequiste columns. The columns can be specified within the function. Other features include:

- Highlighting any missing data essential for citation (e.g. author, name of meeting/conference, etc). The type or level or presentation, and the date of the meeting/conference are considered optional for the purposes of citation.

- The format can be customised via `cite_format` to match preferred referencing style. Note the variable names in the string must exactly match an essential column (e.g. "author" not "Author" or "AUTHOR") 

```{r cite_pre_data, warning=FALSE, message=FALSE, echo=FALSE,results='asis'}
col_pre7 <- which(colnames(data_presentations) %in% c("title"))
col_pre3 <- which(colnames(data_presentations) %in% c("author", "con_org", "con_name"))
col_pre1 <- which(colnames(data_presentations) %in% c("type", "con_country"))

 data_presentations %>%
  knitr::kable(format="html") %>%
  kableExtra::column_spec(col_pre7, width_min="7in") %>%
  kableExtra::column_spec(col_pre3, width_min="3in") %>%
  kableExtra::column_spec(col_pre1, width_min="1.5in") %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F) %>%
  kableExtra::scroll_box(width = "1000px")
```

&nbsp;

```{r cite_pre, warning=FALSE, message=FALSE}
impactr::cite_presentation(data_presentations, cite_format = "author. title. con_org con_name, con_date_range, con_city (con_country).") %>%
  
  dplyr::mutate(citation = gsub("\\]", "", gsub("\\[", "", as.character(citation)))) %>%
  dplyr::select(citation) %>%
  knitr::kable(format="html") %>% kableExtra::kable_styling(bootstrap_options = "striped", full_width = F)
```

&nbsp;
